On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-05-28 LME-MPM.tex
Versión 2
Cambios MP 28-5-2020

\newcommand{\MMS}[1]{
  \textcolor{gree}{{#1}}
}

@@

\title{Enhanced Material Point Method to face
  dynamic problems: Local-maximum entropy approximation and explicit
  predictor-corrector scheme\\
  \PNA{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}}

%% Group authors per affiliation:
\author{
Miguel Molinos$^a$
Pedro Navas$^a$\footnote{Corresponding author: p.navas@upm.es},
Manuel Pastor$^a$
and Miguel Martín Stickle$^a$ 
}
\address{
  $^a$ ETSI Caminos, Canales y Puertos, Universidad Polit\'ectnica de Madrid.\\ c. Prof. Aranguren 3, 28040 Madrid, Spain
}

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an
  alternative to \acrfull{fem} under the large
  deformation regime. However, the simulation of shock waves
  propagation and other high frequency problems is still challenging
  under this approach due the incapability of the standard \acrshort{mpm} time
  integration scheme to filter spurious noises. To overcome this
  limitation in this paper, an Explicit Predictor-Corrector time
  integration scheme has been proposed. Its \textcolor{red}{powerful}
  performance mitigates the presence of spurious oscillations with
  minimal dissipation in high frequency problems. Other source of
  numerical noise in \acrshort{mpm} occurs when the material points
  cross computational grid boundaries, being this issue motivated by
  the lack of smoothness of the interpolation functions. This noise
  results in spurious local variations at the material points, where
  strain-stress fields are computed. This could lead to inaccurate
  solutions as well as aborted simulations in the worst cases. To
  overcome it, this document adopts the \acrfull{lme} approximation
  schemes as a robust substitute of the traditional shape function in
  \acrshort{mpm}. \acrshort{lme} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter
  they are able to behave finite element like or as a meshfree
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements thorough numerical examples.    
\end{abstract}

@@

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}, its popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies.\PNA{One of the main fields where this method is strong is the dynamic one since the original time integration scheme proposed was the \acrfull{fe}~\cite{Sulsky1994}, carried out in a explicit manner. However, in this type or problems, the main instabilities of the original \acrshort{mpm} are even marked.}

\PNA{On the one hand, the main source of instability occurs} when
material points cross cell boundaries. This provoked the development
of other interpolation techniques to overcome this limitation such as
the \acrfull{gimp} method Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrated to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is a voxel-based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the \acrfull{ddmp} method proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases~\cite{DHAKAL2016301}. Therefore \acrshort{ddmp} a large number of particles is needed for convergence, what
makes the method very expensive. In recent years the employment of
spline-lines as shape functions has gain popularity with the introduction of the B-Spline 
\acrshort{mpm} proposed by Roel Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employment of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to the \acrshort{mpm} framework with the \acrfull{ctls}
reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}. Unfortunately, when particles are spread in
a challenging way, the quality of the \acrshort{ctls} approximation decreases locally.

@@

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers have proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing\PNA{somehow} the accuracy.\MMP{One of the first attempts was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} downsized these spurious noises by adding a non-viscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. In a recent publication,
Kularathna \& Soga \cite{Soga_2017} study an implicit treatment of the pressure in MPM algorithm to simulate material incompressibility
avoiding artificial pressure oscillations by applying
Chorin\textquotesingle s
projection. Within the explicit time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort.} In this paper a less time consuming and high efficient explicit predictor-corrector integration method has been
proposed. It consists of an accommodation of the traditional \acrsull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and its good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

@@

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method is composed by three main steps: (i) a
variational recovery process, where particle data is projected to the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention. Three kind of subscript or superscript are used
within paper. The subscript $\Box_p$ is used to define a particle
variable. While the subscript $\Box_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\Box^{\psi}$ involves a
virtual magnitude. For the operators, the convention is : $\dot{\Box}$ and
$\ddot{\Box}$ for the first and second time derivative, $\qued \otimes
\qued$ means the dyadic operator, $\Box \cdot \Box$ and $\Box \colon \Box$ means the
first and second contraction of a tensor, $\Div{\Box}$ denotes the
divergence operator, and finally $\Grad{\Box}$ and $\GradS{\Box}$
denotes the gradient and its symmetric part. Einstein subscripts
convention is adopted therefore repeated index means addition.\PNA{Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are describe in subsection \ref{sec:derivation-mpm},\ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.}

@@

In \acrshort{mpm} the continuum mechanics approach is considered. So on, let
define a continuum \gls{domain}  occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied.\PNA{First (i) the \textit{compatibility
  equation}, where the the strain field \gls{strain} is extracted from \gls{u} is defined as:}
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type \gls{dirichlet-boundary}. An additional consideration over the strain field is
the assumption of inifinitesimal strain, therefore second order terms
in the spatial derivatives can be neglected.\PNA{Therefore, the stress field \gls{stress} will be considered the corresponding conjugate variable for the strain field, being the one} which satisfies (ii) the \textit{conservation of
  momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type \gls{neumann-boundary}.  An aditional component will be (iii) the constitutive equation as a linear
application from $\Re^n$ to $\Re^n$, which relates the
strain tensor with the stress tensor:
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
\PNA{In this research, plane strain Linear Elasticity has been considered. Thus the constitutive tensor, \tens{D}, is the well known linear elastic one.} The final restriction is (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

In order to obtain the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vect{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vect{u}^{\psi} \in
  \mathcal{H}^1 \mid \vect{u}^{\psi} = \vect{0}\ \text{on}\ \Gamma_d \};
\end{equation}
satisfying that the Cauchy sequences are convergent in \gls{domain} as well:
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vect{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty.
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vect{u} \in
\mathcal{H}^1_0$ such that, for $\vect{u}^{\psi} \in
\mathcal{H}^1_0$, the following holds:
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}}.
\end{equation}\\
Thus, equation~\eqref{eq:BalanceMomentum_wf}, together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative}, represents the weak form
formulation of the problem.

In order to obtain a finite set of equations, in contrast with the
\acrshort{fem}, in \acrshort{mpm} a double discretization procedure is
performed. First, the continuum \gls{domain} is discretized with a finite sum of material points (also denominated particles the manuscript). Each material point
represents a part of the discretized domain $\varOmega_p \subset
\varOmega$ with $p = 1,2\ldots ,N_p$ where $N_p$ is the number of
particles. The material point location, $\vec{x}_p$, is defined at the centroid
of each $\Omega_p$ (see figure~\ref{fig:MPM-discretization} for details).
Initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,
$\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$ respectively are assigned to each material point, which also owns the
virtual displacement field $\vect{u}^{\psi}_{p}$. Therefore, employing
the definition of the material integral, where Riemann
integral definition \cite{Riemann_1854} is recovered as an addition of
a finite set of points, and their volumes are interpreted as quadrature weights. Consequently,
individual terms in \eqref{eq:BalanceMomentum_wf} are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \frac{d\vec{v}}{dt} \cdot \vect{u}^{\psi}} =
    \frac{d\vec{v}_{p}}{dt} \cdot \vect{u}^{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \tens{\varepsilon}^{\psi}} =
   \tens{\sigma}_{p}\ \colon \tens{\varepsilon}^{\psi}_p \ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b} \cdot \vect{u}^{\psi} } = 
  \vec{b}_{p} \cdot \vect{u}^{\psi}_p\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vect{u}^{\psi}} = \Integral{2}{\rho\
      \vec{t}^s \cdot \vect{u}^{\psi}} = \vec{t}^s_{p} \cdot \vect{u}^{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
where $h$ is the thickness of the continuum in a 2D case. Following, the aforementioned second discretization procedure appears. A background mesh
composed by a finite set of grid points with coordinates $\vec{x}_I\
, I = 1,2\ldots ,N_n$, is generated, being $N_n$ the number of grid
nodes.\PNA{Spatial derivatives, such as gradients and
divergences, are computed through the support of the background mesh.}

@@

\subsection{\acrshort{mpm} time integration scheme: the Explicit Predictor-Corrector proposal}
%\subsection{Explicit predictor-corrector scheme for \acrshort{mpm}.}
\label{sec:epc-algor-mpm}

>>>>

\newcommand{\MP}[1]{
  \textcolor{green}{{#1}}
}

@@

\title{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}

%% Group authors per affiliation:
\author{
Miguel Molinos$^a\footnote{Corresponding author: m.molinoss@outlook.es}$
Pedro Navas$^a$\footnote{Corresponding author: p.navas@upm.es},
Manuel Pastor$^a$
and Miguel Martín Stickle$^a$ 
}
\address{
  $^a$ ETSI Caminos, Canales y Puertos, Universidad Polit\'ectnica de Madrid.\\ c. Prof. Aranguren 3, 28040 Madrid, Spain
}

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an
  alternative to \acrfull{fem} under the large
  deformation regime. However, the simulation of shock waves
  propagation and \MP{in the large deformation regime} is still challenging
  under this approach due the incapability of the standard \acrshort{mpm} time
  integration scheme to filter spurious noises. \MP{To overcome it ,
    we propose in this paper an explicit Predictor-Corrector time integration scheme}. Its \textcolor{red}{powerful}
  performance mitigates the presence of spurious oscillations with
  minimal dissipation in high frequency problems. Other source of
  numerical noise in \acrshort{mpm} occurs when the material points
  cross computational grid boundaries, being \MP{caused} by
  the lack of smoothness of the interpolation functions. This noise
  results in spurious local variations at the material points, where
  strain-stress fields are computed. This could lead to inaccurate
  solutions as well as aborted simulations in the worst cases. To
  overcome it, \MP{we propose in this work} the \acrfull{lme} approximation
  schemes as a robust substitute of the traditional shape function in
  \acrshort{mpm}. \acrshort{lme} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter
  they are able to behave finite element like or as a meshfree
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements thorough numerical examples.    
\end{abstract}

@@

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}, its popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies.\PNA{One of the main
  fields where this method is strong} \MP{in solid dynamics \PNA{since the original time integration scheme proposed was the \acrfull{fe}~\cite{Sulsky1994}, carried out in a explicit manner. However, in this type or problems, the main instabilities of the original \acrshort{mpm} are even marked.}

\PNA{On the one hand, the main source of instability occurs} when
material points cross cell boundaries. This provoked the development
of other interpolation techniques to overcome this limitation such as
the \acrfull{gimp} method Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrated to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is a voxel-based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed \MP{by} Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the \acrfull{ddmp} method proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases. Therefore \acrshort{ddmp} a large
number of particles is needed for convergence~\cite{DHAKAL2016301}, what makes the method
very expensive. In recent years the employment of spline-lines as shape functions has gain popularity with the introduction of the B-Spline 
\acrshort{mpm} proposed by Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employment of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to the \acrshort{mpm} framework with the \acrfull{ctls}
reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}. Unfortunately, when particles are spread
in \MP{some particular special patterns}, the quality of the \acrshort{ctls} approximation \MP{decrease} locally.

@@

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers have proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing\PNA{somehow} the accuracy.\MMP{One of the first attempts was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016}} \MP{mitigated} \MMP{these spurious noises by adding a non-viscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. In a recent publication,
Kularathna \& Soga \cite{Soga_2017} study an implicit treatment of the pressure in MPM algorithm to simulate material incompressibility
avoiding artificial pressure oscillations by applying
Chorin\textquotesingle s
projection. Within the explicit time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort.} In this paper a less time consuming and high efficient explicit predictor-corrector integration method \MP{is}
proposed. It consists of an accommodation of the traditional \acrsull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

@@

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method \MP{consists of} three main steps: (i) a
variational recovery process, where particle data is projected \MP{onto} the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention. Three kind of subscript or superscript are used
within paper. The subscript $\Box_p$ is used to define a particle
variable. While the subscript $\Box_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\Box^{\psi}$ involves a
virtual magnitude. For the operators, the convention is : $\dot{\Box}$ and
$\ddot{\Box}$ for the first and second time derivative, $\qued \otimes
\qued$ means the dyadic operator, $\Box \cdot \Box$ and $\Box \colon \Box$ means the
first and second contraction of a tensor, $\Div{\Box}$ denotes the
divergence operator, and finally $\Grad{\Box}$ and $\GradS{\Box}$
denotes the gradient and its symmetric part. Einstein subscripts
convention is adopted therefore repeated index means addition.\PNA{Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are describe in subsection \ref{sec:derivation-mpm},\ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.}

@@

In \acrshort{mpm} the continuum mechanics approach is considered. So on, \MP{we will consider a region \gls{domain} }occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied.\MP{First (i) we will relate
  global to local state by the \textit{compatibility equation}, where the the strain field \gls{strain} is extracted from \gls{u} is defined as:}
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type
\gls{dirichlet-boundary}.\MP{We will further assume that strains are
infinitesimal}, therefore second order terms
in the spatial derivatives can be neglected.\MP{From here}, \PNAthe
stress field \gls{stress} will be considered the corresponding
conjugate variable for the strain field, being the one} which
satisfies (ii) the \textit{conservation of \MP{lineal} momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type \gls{neumann-boundary}.  An aditional component will be (iii) the constitutive equation as a linear
application \MP{from $\mathbb{R}^n$ to $\mathbb{R}^n$, which relate stress and strain increments as,}
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
\PNA{In this research, plane strain Linear Elasticity has been
  considered. Thus the constitutive tensor, \tens{D}, is the well
  known linear elastic one.} The final \MP{equation of the set} is (iv) the mass
conservation, which can be obtained by setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

In order to obtain the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vect{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vect{u}^{\psi} \in
  \mathcal{H}^1 \mid \vect{u}^{\psi} = \vect{0}\ \text{on}\ \Gamma_d \};
\end{equation}
\MP{in which} the Cauchy sequences are convergent in \gls{domain} as well:
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vect{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty.
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vect{u} \in
\mathcal{H}^1_0$ such that, for \MP{any} $\vect{u}^{\psi} \in
\mathcal{H}^1_0$,
the following holds:
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}}.
\end{equation}\\
Thus, equation~\eqref{eq:BalanceMomentum_wf}, together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative}, represents the weak form
formulation of the problem.

\MP{Next, we will discretize the set of equations of the mathematical model using a double discretization procedure.} First, the continuum \gls{domain} is discretized with a finite sum of material points (also denominated particles the manuscript). Each material point
represents a part of the discretized domain $\varOmega_p \subset
\varOmega$ with $p = 1,2\ldots ,N_p$ where $N_p$ is the number of
particles. The material point location, $\vec{x}_p$, is defined at the centroid
of each $\Omega_p$ (see figure~\ref{fig:MPM-discretization} for details).
Initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,
$\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$ respectively are assigned to each material point, which also owns the
virtual displacement field $\vect{u}^{\psi}_{p}$. \MP{Regarding numerical integration, the material derivatives are
  evaluated using the Riemann integral definition \cite{Riemann_1854}
  applied to a finite set of points which associated volumes are
  interpreted as quadrature weights.} Consequently,
individual terms in \eqref{eq:BalanceMomentum_wf} are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \frac{d\vec{v}}{dt} \cdot \vect{u}^{\psi}} =
    \frac{d\vec{v}_{p}}{dt} \cdot \vect{u}^{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \tens{\varepsilon}^{\psi}} =
   \tens{\sigma}_{p}\ \colon \tens{\varepsilon}^{\psi}_p \ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b} \cdot \vect{u}^{\psi} } = 
  \vec{b}_{p} \cdot \vect{u}^{\psi}_p\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vect{u}^{\psi}} = \Integral{2}{\rho\
      \vec{t}^s \cdot \vect{u}^{\psi}} = \vec{t}^s_{p} \cdot \vect{u}^{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
where $h$ is the thickness of the continuum in a 2D case. \MP{Next, we
  will proceed with the second step of the discretization. It is based on using a background grid } with coordinates $\vec{x}_I\
, I = 1,2\ldots ,N_n$, is generated, being $N_n$ the number of grid
nodes.\PNA{Spatial derivatives, such as gradients and
divergences, are computed through the support of the background mesh.}

@@

\subsection{\acrshort{mpm} time integration scheme: the Newmark Predictor-Corrector proposal}
%\subsection{Explicit predictor-corrector scheme for \acrshort{mpm}.}
\label{sec:epc-algor-mpm}
